Modeling the Circumnuclear Mediums and Accetion Rates of Quiescent Supermassive Black Holes

### 2014-12-07 ms.tex
Rescaling analytic expression, and adding some references.

The source term in the continuity equation (equation~\ref{eq:drhodt}),
%Should equations be referenced as equation of equation
\begin{align}
  q=\frac{\eta \rhostar}{\th},
\label{eq:q}
\end{align}
represents mass input from stellar winds, which we parameterize in
terms of the fraction $\eta$ of the stellar density $\rhostar$ being
recycled into gas on the Hubble time $\th = 10^{10}$ years (see, e.g.,
\citealt{Ciotti+91}).  In Appendix \ref{app:eta} we estimate that
$\eta \approx 0.11(t_{\star}/t_{\rm h})^{-1.26}$ for a Salpeter IMF,
assuming that all stars are formed in a burst of star formation at a
common age of $t_{\star}$ (equation~[\ref{eq:eta}]).  We adopt
$\eta=0.1$ as fiducial {\bf BDM: take $\eta = 0.1$ instead of 1}, but note that changing $\eta$ is equivalent to
rescaling $\rho$ by the same factor.\footnote{This is {\it not} true
  if cooling is included, since the cooling rate is proportional to
  density {\it squared}.}

@@

\begin{eqnarray}
  \dot{M} &=& \frac{\eta \Menc(\rs)}{\th} \nonumber \\
&\approx&
  \begin{cases}
    1.5 \times 10^{-4} M_{\bullet,8}^{1.76}
    \chi^{-3.8} v_{500}^{-3.8}  \eta_{-1} \Msun \, \pyear& \text{core} \\
    2.6 \times 10^{-4} M_{\bullet,8}^{1.48} 
    \chi^{-2.4} v_{500}^{-2.4}  \eta_{-1} \Msun \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:mdot_analytic}
\end{eqnarray}

@@

\begin{eqnarray}
\eddr&\approx&
  \begin{cases}
    6.6 \times 10^{-5} M_{\bullet,8}^{0.76}
    \chi^{-3.8} v_{500}^{-3.8}  \eta_{-1} \Msun \, \pyear& \text{core} \\
    1.2 \times 10^{-4} \Mbheight^{0.48} 
    \chi^{-2.4} v_{500}^{-2.4}  \eta_{-1} \Msun \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:eddr_analytic}
\end{eqnarray}

The stellar density at the stagnation radius (eq.~[\ref{eq:stag_simple}]) is usefully approximated as
\begin{eqnarray}
  && \rho_{\star}(r_{\rm s}) \approx \nonumber \\
&& \begin{cases}
    1.0\times 10^{-18}M_{\bullet,8}^{-1.2}v_{500}^{2}\chi^{2}\,{\rm g\,cm^{-3}}
& \text{core} \\
    1.1 \times 10^{-18} M_{\bullet,8}^{-1.6}v_{500}^{4}\chi^{4}
      \,{\rm g\,cm^{-3}}  & \text{cusp}, 
  \end{cases}
  \label{eq:rhostarrs}
\end{eqnarray}

The density of gas at the stagnation radius $\rho(r_{\rm s})$ can be roughly estimated by using equations (\ref{eq:stag_simple}) and (\ref{eq:mdot_analytic}) in conjunction with an alternative estimate of the mass accretion rate as the gaseous mass within the stagnation radius divided by the free-fall time $t_{\rm ff}(\rs) = (r_{\rm s}^{3}/GM_{\bullet})^{1/2}$, viz.
\begin{align}
  &\dot{M}\simeq\frac{(4 \pi/3) \rs^3 \rho(\rs)}{t_{\ff}(\rs)},
  \label{eq:mdot_gas}
\end{align}
from which one derives 
\begin{align}
  \rho(\rs)\simeq
  \begin{cases}
    3 \times 10^{-24} \Mbheight^{-0.14} v_{500}^{-1}  \eta \,
    \, {\rm g \, cm^{-3}}& \text{core}\\
    6 \times 10^{-24}  \Mbheight^{-0.56} v_{500}    \eta \,\, {\rm g \,cm^{-3}} & \text{cusp}
  \end{cases}
  \label{eq:rhors}
\end{align}

@@

Our analytic expression for the accretion rate (eq.~[\ref{eq:mdot_analytic}]) can be translated into the growth time of the SMBH $t_{\rm grow} \equiv M_{\bullet}/\dot{M}$,
\begin{eqnarray}
  t_{\rm growth}/t_{\rm h} \approx 
  \begin{cases}
    6.3 M_{\bullet,8}^{-0.9}
    v_{500}^{4}  \eta \Msun \, \pyear& \text{core} \\
    2.9 \Mbheight^{-0.4}
    v_{500}^{-2}  \eta \Msun \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:tgrow_analytic}
\end{eqnarray}
which we have normalized in terms of the Hubble time.

@@

An important ratio controlling the importance of thermal stability is thus
\begin{align}
\left.\frac{q \vw^2/2}{\dot{q}_{\rm cool}}\right|_{\rm r{s}} \simeq
  \begin{cases}
   12\eta^{-1} M_{\bullet,8}^{-0.9}v_{500}^{7.2}\chi^{7.2}  &, \text{core}\\
   4\eta^{-1} M_{\bullet,8}^{-0.5}v_{500}^{5.8}\chi^{5.8}  &, \text{cusp},     
  \end{cases}
  \label{eq:cooling2}
\end{align}
where we have used equations (\ref{eq:rhostarrs}) and
(\ref{eq:rhors}).  We thus see that we require $v_{\rm w} \gtrsim $
300 km s$^{-1}$ for stability, with a weak sensitivity on $\eta$ and
$M_{\bullet}$.  {\bf AG:I flipped the ratio to be consistent with the
  ratio of timescales given below (NB here we are comparing heating
  and cooling rates as opposed to timescales). Modulo factors of order
  unity I find no inconsistency. We should probably only keep one of
  these ratios. Note I have adopted the new convention of Gamma=0.8 is
  a cusp galaxy and Gamma=0.1 is a core galaxy.}

@@

Using equations (\ref{eq:Tanalytic}), (\ref{eq:stag_simple}), (\ref{eq:rhors}), (\ref{eq:cooling}), (\ref{eq:cooling2}), we can estimate the ratio of the cooling timescale to the free-fall time $\sim r/\sigma$ at the stagnation radius
\begin{align}
  \left.\frac{t_{\rm cool}}{t_{\rm ff}}\right|_{r_{\rm s}}\simeq
  \begin{cases}
    11 \Mbheight^{-0.8} \eta^{-1}(\chi v_{500})^{7.2}, & \text{core}\\
    6  \Mbheight^{-0.5} \eta^{-1}(\chi v_{500})^{5.8} ,   & \text{cusp}
  \end{cases}
  \label{eq:coolingratio}
\end{align}
We thus see that $t_{\rm cool} \lesssim 10 t_{\rm ff}$ at the
stagnation radius for low $v_{\rm w} \lesssim 250$ km s$^{-1}$ if
$\eta \sim 0.1$, but even for higher $v_{\rm w}$ if $\eta \sim 1$.

@@

  \section{Analytic model for dependence of wind mass loss rate on stellar age}
  \label{app:eta}
  \input{eta}

>>>>

The source term in the continuity equation (equation~\ref{eq:drhodt}),
\begin{align}
  q=\frac{\eta \rhostar}{\th},
\label{eq:q}
\end{align}
represents mass input from stellar winds, which we parameterize in
terms of the fraction $\eta$ of the stellar density $\rhostar$ being
recycled into gas on the Hubble time $\th = 1.4 \times 10^{10}$
years. $\eta\simeq 0.02 \left(\frac{t}{th}\right)^{-1.3}$ 
(see, e.g., \citealt{Ciotti+91})\footnote{Note that 
\citealt{Ciotti+91} gives the overall mass return rate from evolved
stars as a function of B-band luminosity. We give the mass return rate
per unit volume. However, our expressions are equivalent.}

@@

\begin{eqnarray}
  \dot{M} &=& \frac{\eta \Menc(\rs)}{\th} \nonumber \\
&\approx&
  \begin{cases}
    2.9 \times 10^{-5} M_{\bullet,8}^{1.76}
    \chi^{-3.8} v_{500}^{-3.8}  \eta_{0.02} \Msun \, \pyear& \text{core} \\
    5.3 \times 10^{-5} M_{\bullet,8}^{1.48} 
    \chi^{-2.4} v_{500}^{-2.4}  \eta_{0.02} \Msun \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:mdot_analytic}
\end{eqnarray}

@@

\begin{eqnarray}
\eddr&\approx&
  \begin{cases}
    1.3 \times 10^{-5} M_{\bullet,8}^{0.76}
    \chi^{-3.8} v_{500}^{-3.8}  \eta_{0.02}   & \text{core} \\
    2.4 \times 10^{-5} \Mbheight^{0.48} 
    \chi^{-2.4} v_{500}^{-2.4}  \eta_{0.02}   & \text{cusp}, 
  \end{cases}
  \label{eq:eddr_analytic}
\end{eqnarray}

The stellar density at the stagnation radius (eq.~[\ref{eq:stag_simple}]) is usefully approximated as
\begin{eqnarray}
  && \rho_{\star}(r_{\rm s}) \approx \nonumber \\
  && \begin{cases}
    9.4 \times 10^{-19}M_{\bullet,8}^{-1.2}v_{500}^{2.2}\chi^{2.2}\,{\rm g\,cm^{-3}}
    & \text{core} \\
    1.1 \times 10^{-18} M_{\bullet,8}^{-1.5}v_{500}^{3.6}\chi^{3.6}
    \,{\rm g\,cm^{-3}}  & \text{cusp}, 
  \end{cases}
  \label{eq:rhostarrs}
\end{eqnarray}

The density of gas at the stagnation radius $\rho(r_{\rm s})$ can be roughly estimated by using equations (\ref{eq:stag_simple}) and (\ref{eq:mdot_analytic}) in conjunction with an alternative estimate of the mass accretion rate as the gaseous mass within the stagnation radius divided by the free-fall time $t_{\rm ff}(\rs) = (r_{\rm s}^{3}/GM_{\bullet})^{1/2}$, viz.
\begin{align}
  &\dot{M}\simeq\frac{(4 \pi/3) \rs^3 \rho(\rs)}{t_{\ff}(\rs)},
  \label{eq:mdot_gas}
\end{align}
from which one derives 
\begin{align}
  \rho(\rs)\simeq
  \begin{cases}
    5 \times 10^{-26} \Mbheight^{-0.2} v_{500}^{-0.8}  \eta_{0.02} \,
    \, {\rm g \, cm^{-3}}& \text{core}\\
    9 \times 10^{-26}  \Mbheight^{-0.5} v_{500}^{0.6}  \eta_{0.02} \,\, {\rm g \,cm^{-3}} & \text{cusp}
  \end{cases}
  \label{eq:rhors}
\end{align}

@@

Our analytic expression for the accretion rate (eq.~[\ref{eq:mdot_analytic}]) can be translated into the growth time of the SMBH $t_{\rm grow} \equiv M_{\bullet}/\dot{M}$,
\begin{eqnarray}
  t_{\rm growth}/t_{\rm h} \approx 
  \begin{cases}
    250 M_{\bullet,8}^{-0.76}
    v_{500}^{3.8} \chi^{3.8}  \eta_{0.02}  & \text{core} \\
    140 \Mbheight^{-0.48}
    v_{500}^{2.4} \chi^{2.4}  \eta_{0.02}  & \text{cusp}, 
  \end{cases}
  \label{eq:tgrow_analytic}
\end{eqnarray}
which we have normalized in terms of the Hubble time.

@@

An important ratio controlling the importance of thermal stability is thus
\begin{align}
\left.\frac{q \vw^2/2}{\dot{q}_{\rm cool}}\right|_{\rm r{s}} \simeq
  \begin{cases}
   590 \eta_{0.02}^{-1} M_{\bullet,8}^{-0.76}v_{500}^{7.2}\chi^{7.2}  &, \text{core}\\
   210 \eta_{0.02}^{-1} M_{\bullet,8}^{-0.48}v_{500}^{5.8}\chi^{5.8}  &, \text{cusp},     
  \end{cases}
  \label{eq:cooling2}
\end{align}
where we have used equations (\ref{eq:rhostarrs}) and
(\ref{eq:rhors}).  We thus see that we require $v_{\rm w} \gtrsim $
300 km s$^{-1}$ for stability, with a weak sensitivity on $\eta$ and
$M_{\bullet}$.  {\bf AG:I flipped the ratio to be consistent with the
  ratio of timescales given below (NB here we are comparing heating
  and cooling rates as opposed to timescales). Modulo factors of order
  unity I find no inconsistency. We should probably only keep one of
  these ratios. Note I have adopted the new convention of Gamma=0.8 is
  a cusp galaxy and Gamma=0.1 is a core galaxy.}

@@

Using equations (\ref{eq:Tanalytic}), (\ref{eq:stag_simple}), (\ref{eq:rhors}), (\ref{eq:cooling}), (\ref{eq:cooling2}), we can estimate the ratio of the cooling timescale to the free-fall time $\sim r/\sigma$ at the stagnation radius
\begin{align}
  \left.\frac{t_{\rm cool}}{t_{\rm ff}}\right|_{r_{\rm s}}\simeq
  \begin{cases}
    560 \Mbheight^{-0.76} \eta_{0.02}^{-1}(\chi v_{500})^{7.2}, & \text{core}\\
    310  \Mbheight^{-0.48} \eta_{0.02}^{-1}(\chi v_{500})^{5.8} ,   & \text{cusp}
  \end{cases}
  \label{eq:coolingratio}
\end{align}
We thus see that $t_{\rm cool} \lesssim 10 t_{\rm ff}$ at the
stagnation radius for low $v_{\rm w} \lesssim 250$ km s$^{-1}$ if
$\eta \sim 0.1$, but even for higher $v_{\rm w}$ if $\eta \sim 1$.

@@

  % \section{Analytic model for dependence of wind mass loss rate on stellar age}
  % \label{app:eta}
%  \input{eta}
